Reasoning about a Capability Machine with Local Capabilities

### 2017-06-23 popl-2018/paper.tex
related

Our logical relation is a unary, step-indexed Kripke logical relation
with recursive
worlds~\citep{Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
capability safety in a high-level JavaScript-like lambda calculus. Our
Fundamental Theorem is similar to theirs and expresses
capability safety of the capability machine. Because we are not interested in
externally observable side-effects (like console output or memory access
traces), we do not require their notion of effect parametricity. Our logical
relation uses several ideas from previous work, like Kripke worlds with regions
containing state transition systems~\citep{Ahmed:popl09}, public/private future
worlds~\citep{Dreyer:jfp12} (see Section~\ref{sec:discussion} for a discussion), and
biorthogonality~\citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.
Swasey et. al.\ have recently developed a \emph{logic}, OCPL,  for verification of
object capability patterns\citep{swasey:2017}. The logic is based on
Iris~\cite{iris,iris2,iris3}, a state of the art higher-order
concurrent separation logic and is formalized in Coq, building on the
Iris Proof Mode for Coq~\cite{ipm}.  
XXX HERE

>>>>

Our logical relation is a unary, step-indexed Kripke logical relation
with recursive
worlds~\citep{Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
capability safety in a high-level JavaScript-like lambda calculus. Our
Fundamental Theorem is similar to theirs and expresses
capability safety of the capability machine. Because we are not interested in
externally observable side-effects (like console output or memory access
traces), we do not require their notion of effect parametricity. Our logical
relation uses several ideas from previous work, like Kripke worlds with regions
containing state transition systems~\citep{Ahmed:popl09}, public/private future
worlds~\citep{Dreyer:jfp12} (see Section~\ref{sec:discussion} for a discussion), and
biorthogonality~\citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.
Swasey et. al.\ have recently developed a \emph{logic}, OCPL,  for verification of
object capability patterns~\citep{swasey:2017}. The logic is based on
Iris~\citep{iris,iris2,iris3}, a state of the art higher-order
concurrent separation logic and is formalized in Coq, building on the
Iris Proof Mode for Coq~\citep{ipm}.  OCPL gives a more abstract and
modular way of proving capability safety for a lambda-calculus (with concurrency)
compared to the earlier work by \citet{Devriese:2016ObjCap}. 
In the future we would also like to investigate a new program logic
for reasoning about capability safety for our capability machine
model.  We think Iris would also be a natural starting point for such
an endaveour, since Iris is really a framework, which can be instantiated to
different programming languages. OCPL was able to leverage existing
Iris specifications for the high-level language; for our capability
machine model, however, it would be necessary  to devise new kinds of
specifications for our low-level programs
with unstructured control-flow. It is likely that we could get
inspiration from earlier work on logics for assembly programming
languages, such as XCAP~\citep{xcap}.
